---
title: "Yuma_PCA_Analysis"
author: "Jenna Honan"
date: "6/29/2022"
output: html_document
---
# An Analysis of PC1 and PC2 Compared to Yuma Health Outcomes

Note: The following scripts must be run prior to completing this analysis: Data_Clean.Rmd, Yuma_Metals_Analysis.Rmd, Demographics.Rmd, yuma-metals-20210821.Rmd. This will provide the accurate dataframes used in the following code.

## Data Descriptions
```{r desc}
## Required libraries
library(DataExplorer)
library(ggplot2)
library(ggpubr)
library(gridExtra)
library(dplyr)
library(tidyverse)
library(scales)

Metals_PCA <- merge(Metals, metals.robpca, by = "ID", all.x = TRUE)
Metals_PCA <- Metals_PCA %>% dplyr::select(-Mn55.y, -Cu65.y, -Cd111.y, -Hg202.y, -Pb208.y, -U238.y, -Site.y)
colnames(Metals_PCA)[colnames(Metals_PCA) == 'Site.x'] <- 'Site'

#create_report(Metals) (Remove the # when you actually want to run - takes too much processing power otherwise.)
```


## Data Visualization - Exploratory


### Tables and Observations for Each Metal
```{r observations_envt_samples}


Metals_PCA %>% 
  group_by(Site) %>% 
  summarize(number_NAs_PC1=sum(is.na(PC1)),
            number_NAs_PC2=sum(is.na(PC2)))

#Not sure why I made this:
#Making a table re: PC values
#library(EnvStats)
#knitr::kable(Metals %>%
#               group_by(Site) %>% 
#               summarize(n_PC1=sum(!is.na(PC1)),
#                         geomean_PC1=round(exp(mean(PC1, na.rm = T)), digits = 2),
#                         geosd_PC1=round(exp(sd(PC1, na.rm = T)), digits = 2),
#                         n_PC2=sum(!is.na(PC2)),
#                         geomean_PC2=round(exp(mean(PC2, na.rm = T)), digits = 2),
#                         geosd_PC2=round(exp(sd(PC2, na.rm = T)), digits = 2)))

#create_report(data.melt) (Only run if you need to.)
```

### Distribution of PCA Data
```{r metal_distributions}
hist1 <- hist(Metals_PCA$PC1)

hist2 <- hist(Metals_PCA$PC2)

plot(hist1, col=rgb(0,0,1,1/4), ylim=c(0,100), main="Distributions of PCs", xlab = "")  # first histogram,  xlim=c(0,10)
plot(hist2, col=rgb(1,0,0,1/4), ylim=c(0,100), add=T) # second histogram ,  xlim=c(0,10),
legend("topright", c("PC1", "PC2"), fill=c(rgb(0,0,1,1/4), rgb(1,0,0,1/4)))
box()

```

### Plotting Hormone Levels Against Metal Concentrations
```{r metals_hormones}
#Cortisol against PCA results
PC1_Cortisol <- ggplot(Metals_PCA, aes(x=PC1, y=Cortisol))+geom_point()
PC2_Cortisol <- ggplot(Metals_PCA, aes(x=PC2, y=Cortisol))+geom_point()

grid.arrange(PC1_Cortisol, PC2_Cortisol)

#TSH against PCA results
PC1_TSH <- ggplot(Metals_PCA, aes(x=PC1, y=log10(TSH)))+geom_point()
PC2_TSH <- ggplot(Metals_PCA, aes(x=PC2, y=log10(TSH)))+geom_point()

grid.arrange(PC1_TSH, PC2_TSH)

#Free T4 against PCA results
PC1_fT4 <- ggplot(Metals_PCA, aes(x=PC1, y=fT4))+geom_point()
PC2_fT4 <- ggplot(Metals_PCA, aes(x=PC2, y=fT4))+geom_point()

grid.arrange(PC1_fT4, PC2_fT4)
```

### PCA Stuff from Dean

Pb isotopes were removed because they were all heavily collinear. 
Perhaps so much so that the isotopes were calculated based on one reference one.
Kept Pb208.

The classic vs. robust PCAs were different enough that Dean felt like keeping the robust version was worth it. Even after logging the data, they are still skewed.

PC1 is all positively correlated, which shows that it is a good analysis of overall exposure.

PC2 shows which ones are correlated in the same +ve or -ve directions.

### PCA by Gender
```{r PCA_Demo}
Gender_PC1 <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Gender)), aes(y=PC1, x=as.factor(Gender), fill=as.factor(Gender))) +
 geom_boxplot() +
  geom_jitter(alpha=0.15)+
 scale_x_discrete(labels = c("1" = "Male", "2" = "Female")) +
 theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000")) +
 ylab("PC1") +
 xlab("")

Gender_PC2 <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Gender)), aes(y=PC2, x=as.factor(Gender), fill=as.factor(Gender))) +
 geom_boxplot() +
  geom_jitter(alpha=0.15)+
 scale_x_discrete(labels = c("1" = "Male", "2" = "Female")) +
 theme(legend.position = "none",
       panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000")) +
 ylab("PC2") +
 xlab("")

ggarrange(Gender_PC1, Gender_PC2, ncol = 2, nrow = 1, common.legend = FALSE)
```

### PCA by Farmer Status
```{r PCA_Demo}
counts <- Metals %>% 
  group_by(Farmer) %>% 
  summarise(count = n())

Farmer_PC1 <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Farmer)), aes(y=PC1, x=as.factor(Farmer), fill=as.factor(Farmer))) +
 geom_boxplot() +
  geom_jitter(alpha=0.15)+
 scale_x_discrete(labels = c("1" = "Farmer", "2" = "Non-Farmer")) +
 theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"),
       text = element_text(size = 15)) +
 ylab("PC1") +
 xlab("")

Farmer_PC2 <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Farmer)), aes(y=PC2, x=as.factor(Farmer), fill=as.factor(Farmer))) +
 geom_boxplot() +
  geom_jitter(alpha=0.15)+
 scale_x_discrete(labels = c("1" = "Farmer", "2" = "Non-Farmer")) +
 theme(legend.position = "none",
       panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"),
       text = element_text(size = 15)) +
 ylab("PC2") +
 xlab("")

ggarrange(Farmer_PC1, Farmer_PC2, ncol = 2, nrow = 1, common.legend = FALSE)
```

### PCA by BMI
```{r PCA_Demo}
aa <- ggplot(Metals_PCA, aes(x=PC1, y=log(bmi)))+
  geom_point()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("PC1")+
  ylab("Log BMI (Measured at Sampling Time)")

bb <- ggplot(Metals_PCA, aes(x=PC2, y=log(bmi)))+
  geom_point()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("PC2")+
  ylab("")

ggpubr::ggarrange(aa,bb,
                  ncol = 2, nrow = 1,
                  common.legend = T)

a <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Gender)), aes(x=PC1, y=log(bmi), color=as.factor(Gender))) +
 geom_point() +
 scale_color_discrete(name = "Gender", labels = c("1" = "Male", "2" = "Female")) +
 theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000")) +
 xlab("PC1") +
 ylab("Log BMI")

b <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Gender)), aes(x=PC2, y=log(bmi), color=as.factor(Gender))) +
 geom_point() +
 scale_color_discrete(name = "Gender", labels = c("1" = "Male", "2" = "Female")) +
 theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000")) +
 xlab("PC2") +
 ylab("")

ggarrange(a, b, ncol = 2, nrow = 1, common.legend = TRUE)

ggplot(Metals_PCA, aes(x=PC1, y=bmi))+
    geom_point()+
    geom_smooth(method = lm, se=F)+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    xlab("PC1")+
    ylab("BMI")
```

### PCA by age
```{r PCA_Demo}
e <- ggplot(Metals_PCA, aes(y=PC1, x=age))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("Age")+
  ylab("PC1")

f <- ggplot(Metals_PCA, aes(y=PC2, x=age))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("Age")+
  ylab("PC2")

ggpubr::ggarrange(e,f,
                  ncol = 1, nrow = 2,
                  common.legend = T)

c <- ggplot(Metals_PCA, aes(y=PC1, x=age, color=as.factor(Gender)))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("Age")+
  ylab("PC1")

d <- ggplot(Metals_PCA, aes(y=PC2, x=age, color=as.factor(Gender)))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  scale_color_discrete(name="Gender", labels=c("Male", "Female", NA))+
  labs(Color='Gender')+
  xlab("Age")+
  ylab("PC2")

ggpubr::ggarrange(c,d,
                  ncol = 2, nrow = 1,
                  common.legend = T)


PC1_ClO4 <- ggplot(Metals_PCA, aes(y=PC1, x=as.numeric(Result_P_ng_mL)))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC1")

PC2_ClO4 <- ggplot(Metals_PCA, aes(y=PC2, x=as.numeric(Result_P_ng_mL)))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("Log10 [Perchlorate]")+
  ylab("PC2")

ggpubr::ggarrange(PC1_ClO4, PC2_ClO4,
                  ncol = 1, nrow = 2)
```

``` {r other_outcomes}
o <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Dental_health)), aes(y=PC1, x=as.factor(Dental_health), fill=as.factor(Dental_health)))+
         geom_boxplot()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC1")

p <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Dental_health)), aes(y=PC2, x=as.factor(Dental_health), fill=as.factor(Dental_health)))+
         geom_boxplot()+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC2")

ggpubr::ggarrange(o,p,
                  ncol = 2, nrow = 1,
                  common.legend = T)

q <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$arthritis)), aes(y=PC1, x=as.factor(arthritis), fill=as.factor(arthritis)))+
         geom_boxplot()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC1")

r <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$arthritis)), aes(y=PC2, x=as.factor(arthritis), fill=as.factor(arthritis)))+
         geom_boxplot()+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC2")

ggpubr::ggarrange(q,r,
                  ncol = 2, nrow = 1,
                  common.legend = T)

s <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$gout)), aes(y=PC1, x=as.factor(gout), fill=as.factor(gout)))+
         geom_boxplot()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC1")

t <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$gout)), aes(y=PC2, x=as.factor(gout), fill=as.factor(gout)))+
         geom_boxplot()+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC2")

ggpubr::ggarrange(s,t,
                  ncol = 2, nrow = 1,
                  common.legend = T)

u <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$chf)), aes(y=PC1, x=as.factor(chf), fill=as.factor(chf)))+
         geom_boxplot()+
  theme(legend.position = "none",
        panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC1")

v <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$chf)), aes(y=PC2, x=as.factor(chf), fill=as.factor(chf)))+
         geom_boxplot()+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.text.x=element_text(angle=60, hjust=1),
        axis.line = element_line(color="#000000"))+
  xlab("")+
  ylab("PC2")

ggpubr::ggarrange(u, v,
                  ncol = 2, nrow = 1,
                  common.legend = T)


```

```{r MLMs}
#https://stats.stackexchange.com/questions/565567/visualizing-multivariate-multiple-regression-of-continuous-data-in-r
```

```{r demographics_PCA}
#Looking to see if any demographic data influences the PCA

##Gender
PC1_Gender <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Gender)))+
  geom_jitter(aes(color=as.factor(Gender)))+
  scale_color_discrete(name="Gender",
                       labels=c("Male", "Female"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Gender <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Gender)))+
  geom_jitter(aes(color=as.factor(Gender)))+
  scale_color_discrete(name="Gender",
                       labels=c("Male", "Female"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(PC1_Gender)

#grid.arrange(PC1_Gender, PC2_Gender, ncol=2, nrow=1) - simple version

grid.arrange(arrangeGrob(PC1_Gender + theme(legend.position="none"),
                         PC2_Gender + theme(legend.position="none"),
                         nrow=2),
             mylegend, nrow=1)

##Gender and Age
PC1_age_gender <- ggplot(Metals_PCA, aes(y=PC1, x=age))+
  geom_jitter(aes(color=as.factor(Gender)))+
  scale_color_discrete(name="Gender",
                       labels=c("Male", "Female"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_age_gender <- ggplot(Metals_PCA, aes(y=PC2, x=age))+
  geom_jitter(aes(color=as.factor(Gender)))+
  scale_color_discrete(name="Gender",
                       labels=c("Male", "Female"))+
  xlab("Age")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        panel.spacing = unit(0.75, "cm"))

grid.arrange(arrangeGrob(PC1_age_gender + theme(legend.position="none"),
                         PC2_age_gender + theme(legend.position="none"),
                         nrow=2),
             mylegend, nrow=1)

PCA_Medians <- Metals_PCA %>%
  group_by(Pesticides) %>%
  summarize(median_PC1=round(median(PC1, na.rm = T), digits = 2),
            median_PC2=round(median(PC2, na.rm = T), digits = 2))

#ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Pesticides)))+
#  geom_jitter(aes(color=as.factor(Pesticides)))+
#  geom_hline(data=PCA_Medians, aes(yintercept=median_PC1, color=as.factor(Pesticides)))+
#  scale_color_discrete(name="Pesticide Home Use",
#                       labels=c("Yes", "No", "Don't Know"))+
#  xlab("")+
#  theme(panel.background = element_rect(fill="#ffffff"),
#        axis.line = element_line(color="#000000"),
#        text = element_text(size = 15),
#        axis.text.x = element_blank(),
#        axis.ticks.x = element_blank(),
#        panel.spacing = unit(0.75, "cm"))+
#  ggtitle("PC Values (With Median) by \nSelf-Reported Home Pesticide Use")

##Pesticides
PC1_Pesticides <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Pesticides)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Pesticides)), alpha=0.3)+
  scale_color_discrete(name="Pesticides in Workplace",
                       labels=c("Yes", "No", "Don't Know"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Pesticides <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Pesticides)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Pesticides)), alpha=0.3)+
  scale_color_discrete(name="Pesticides in Workplace",
                       labels=c("Yes", "No", "Don't Know"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend2<-g_legend(PC1_Pesticides)

grid.arrange(arrangeGrob(PC1_Pesticides + theme(legend.position="none"),
                         PC2_Pesticides + theme(legend.position="none"),
                         nrow=2),
             mylegend2, nrow=1)

pesticide_sig <- Metals_PCA %>%
  filter(Pesticides!=3)

t.test(PC1~as.factor(Pesticides), data=pesticide_sig)

t.test(PC2~as.factor(Pesticides), data=pesticide_sig)


##Water Source
#Water 1: City
#Water 2: Well
#Water 3: Bottled
#Water 4: Other
#Water 5: Mixed Sources
PC1_Water <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Water_all)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Water_all)), alpha=0.3)+
  scale_color_discrete(name="Water Source",
                       labels=c("City", "Well", "Bottled", "Other", "Mixed Sources"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Water <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Water_all)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Water_all)), alpha=0.3)+
   scale_color_discrete(name="Water Source",
                       labels=c("City", "Well", "Bottled", "Other", "Mixed Sources"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend3<-g_legend(PC1_Water)

grid.arrange(arrangeGrob(PC1_Water + theme(legend.position="none"),
                         PC2_Water + theme(legend.position="none"),
                         nrow=2),
             mylegend3, nrow=1)

PC1_by_Water <- aov(PC1~as.factor(Water_all), data=Metals_PCA)
summary(PC1_by_Water)

PC2_by_Water <- aov(PC2~as.factor(Water_all), data=Metals_PCA)
summary(PC2_by_Water)

##Recruitment Site
#1 = CSF
#2 = RCBH
#3 = YRMC
PC1_Site <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Site)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Site)), alpha=0.3)+
  scale_color_discrete(name="Recruitment Site",
                       labels=c("CSF", "RCBH", "YRMC"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Site <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Site)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Site)), alpha=0.3)+
  scale_color_discrete(name="Recruitment Site",
                       labels=c("CSF", "RCBH", "YRMC"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend4<-g_legend(PC1_Site)

grid.arrange(arrangeGrob(PC1_Site + theme(legend.position="none"),
                         PC2_Site + theme(legend.position="none"),
                         nrow=2),
             mylegend4, nrow=1)

PC1_by_Site <- aov(PC1~as.factor(Site), data=Metals_PCA)
summary(PC1_by_Site)
pairwise.t.test(Metals_PCA$PC1, as.factor(Metals_PCA$Site), p.adj="bonf")

PC2_by_Site <- aov(PC2~as.factor(Site), data=Metals_PCA)
summary(PC2_by_Site)
pairwise.t.test(Metals_PCA$PC2, as.factor(Metals_PCA$Site), p.adj="bonf")

##Smoking
#From self-report
#0: Never smoker
#1: Former Smoker
#2: Light/Somedays Smoker
#3: Every day smoker
PC1_Smoking <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Smoking_Status)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Smoking_Status)), alpha=0.3)+
    scale_color_discrete(name="Smoking Status",
                       labels=c("Never", "Former", "Light/Occasional", "HeavyDaily"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Smoking <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Smoking_Status)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Smoking_Status)), alpha=0.3)+
    scale_color_discrete(name="Smoking Status",
                       labels=c("Never", "Former", "Light/Occasional", "HeavyDaily"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend5<-g_legend(PC1_Smoking)

grid.arrange(arrangeGrob(PC1_Smoking + theme(legend.position="none"),
                         PC2_Smoking + theme(legend.position="none"),
                         nrow=2),
             mylegend5, nrow=1)

PC1_by_Smoking <- aov(PC1~as.factor(Smoking_Status), data=subset(Metals_PCA, !is.na(Metals_PCA$Smoking_Status)))
summary(PC1_by_Smoking)


PC2_by_Smoking <- aov(PC2~as.factor(Smoking_Status), data=subset(Metals_PCA, !is.na(Metals_PCA$Smoking_Status)))
summary(PC2_by_Smoking)


##Ethnicity
#1: Hispanic
#2: Non-Hispanic
PC1_Ethnicity <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Ethnicity)), aes(y=PC1, x=as.factor(Ethnicity)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Ethnicity)), alpha=0.3)+
  scale_color_discrete(name="Ethnicity",
                       labels=c("Hispanic", "Non-Hispanic"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Ethnicity <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Ethnicity)), aes(y=PC2, x=as.factor(Ethnicity)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Ethnicity)), alpha=0.3)+
  scale_color_discrete(name="Ethnicity",
                       labels=c("Hispanic", "Non-Hispanic"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend6<-g_legend(PC1_Ethnicity)

grid.arrange(arrangeGrob(PC1_Ethnicity + theme(legend.position="none"),
                         PC2_Ethnicity + theme(legend.position="none"),
                         nrow=2),
             mylegend6, nrow=1)

t.test(PC1~as.factor(Ethnicity), data=Metals_PCA)

t.test(PC2~as.factor(Ethnicity), data=Metals_PCA)

##Time Lived in Yuma

#1: Less than 6 months
#2: 6 months to 1 year
#3: 1 to 3 years
#4: 3 to 5 years
#5: More than 5 years

Metals_PCA$Lived_in_Yuma <- factor(Metals_PCA$Lived_in_Yuma, levels = 1:5)

PC1_Yuma_Time <- ggplot(Metals_PCA, aes(y=PC1, x=as.factor(Lived_in_Yuma))) +
 geom_boxplot() +
 geom_jitter(aes(color=as.factor(Lived_in_Yuma)), alpha=0.6) +
 scale_color_discrete(name="Time Lived in Yuma",
                       labels=c("1"="Less than 6 months", 
                                "2"="6 months to 1 year", 
                                "3"="1 to 3 years", 
                                "4"="3 to 5 years", 
                                "5"="More than 5 years"),
                       drop = FALSE) +
  scale_x_discrete(drop=FALSE)+
 xlab("") +
 theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Yuma_Time <- ggplot(Metals_PCA, aes(y=PC2, x=as.factor(Lived_in_Yuma))) +
 geom_boxplot() +
 geom_jitter(aes(color=as.factor(Lived_in_Yuma)), alpha=0.6) +
 scale_color_discrete(name="Time Lived in Yuma",
                       labels=c("1"="Less than 6 months", 
                                "2"="6 months to 1 year", 
                                "3"="1 to 3 years", 
                                "4"="3 to 5 years", 
                                "5"="More than 5 years"),
                       drop = FALSE) +
  scale_x_discrete(drop=FALSE)+
 xlab("") +
 theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

#Extract the legend
mylegend7 <- g_legend(PC1_Yuma_Time)

#Arrange the plots and the legend
grid.arrange(arrangeGrob(PC1_Yuma_Time + theme(legend.position="none"),
                         PC2_Yuma_Time + theme(legend.position="none"),
                         nrow=2),
             mylegend7, nrow=1)

##Income
#1:0-4,999 23 6.9%
#2:5k-9,999 33 9.9%
#3:10k-14,999 47 14.2%
#4:15k-19,999 34 10.2%
#5:20k-29,999 63 19.0%
#6:30k-39,999 34 10.2%
#7:40k-49,999 16 4.8%
#8:50k-59,999 9 2.7%
#9:60k-69,999 10 3.0%
#10:>70k 25 7.5%
#11:Don't know 15 4.5%
#12:Refused 14 4.2%

PC1_Income <- ggplot(subset(Metals_PCA ,!is.na(Metals_PCA$Income)), aes(y=PC1, x=as.factor(Income)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Income)), alpha=0.3)+
    scale_color_discrete(name="Income Group",
                       labels=c("0-4,999", "5k-9,999", "10k-14,999", "15k-19,999", "20k-29,999", "30k-39,999", "40k-49,999", "50k-59,999", "60k-69,999", "70k+", "Don't know", "Refused"),
                       guide=F)+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        #axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Income <- ggplot(subset(Metals_PCA ,!is.na(Metals_PCA$Income)), aes(y=PC2, x=as.factor(Income)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Income)), alpha=0.3)+
    scale_color_discrete(name="Income Group",
                       labels=c("0-4,999", "5k-9,999", "10k-14,999", "15k-19,999", "20k-29,999", "30k-39,999", "40k-49,999", "50k-59,999", "60k-69,999", "70k+", "Don't know", "Refused"),
                       guide=F)+
  scale_x_discrete(name="\nIncome Group",
                       labels=c("0-4,999", "5k-9,999", "10k-14,999", "15k-19,999", "20k-29,999", "30k-39,999", "40k-49,999", "50k-59,999", "60k-69,999", "70k+", "Don't know", "Refused"))+
  xlab("Income Group")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_text(angle = 45, hjust = 1),
        #axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend5<-g_legend(PC1_Smoking)

grid.arrange(PC1_Income, PC2_Income, nrow=2)

PC1_by_Income <- aov(PC1~as.factor(Income), data=subset(Metals_PCA, !is.na(Metals_PCA$Income)))
summary(PC1_by_Income)

PC2_by_Income <- aov(PC2~as.factor(Income), data=subset(Metals_PCA, !is.na(Metals_PCA$Income)))
summary(PC2_by_Income)


##Poverty
#1: Below Pov Line
#0: Not Below Pov Line
PC1_Pov <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Below_Pov)), aes(y=PC1, x=as.factor(Below_Pov)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Below_Pov)), alpha=0.3)+
  scale_color_discrete(name="Poverty Line",
                       labels=c("Above", "Below"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

PC2_Pov <- ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Below_Pov)), aes(y=PC2, x=as.factor(Below_Pov)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(Below_Pov)), alpha=0.3)+
  scale_color_discrete(name="Poverty Line",
                       labels=c("Above", "Below"))+
  xlab("")+
  theme(panel.background = element_rect(fill="#ffffff"),
        axis.line = element_line(color="#000000"),
        text = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.75, "cm"))

mylegend6<-g_legend(PC1_Pov)

grid.arrange(arrangeGrob(PC1_Pov + theme(legend.position="none"),
                         PC2_Pov + theme(legend.position="none"),
                         nrow=2),
             mylegend6, nrow=1)

t.test(PC1~as.factor(Below_Pov), data=Metals_PCA)

t.test(PC2~as.factor(Below_Pov), data=Metals_PCA)


ggplot(Metals_PCA, aes(x=PC1, y=Num_children))+
    geom_point()+
    geom_smooth(method = lm, se=F)+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    xlab("PC1")+
    ylab("Number of Children")

ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Num_children)), aes(y=PC1, x=as.factor(Num_children)))+
    geom_boxplot()+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    ylab("PC1")+
    xlab("Number of Children")

ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Residence_type)), aes(y=PC1, x=as.factor(Residence_type)))+
    geom_boxplot()+
  scale_x_discrete(labels = c("House", "Apt", "Trailer/Mobile", "Group", "Other"))+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    ylab("PC1")+
    xlab("Residence Type")

ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Health_compare)), aes(y=PC1, x=as.factor(Health_compare)))+
    geom_boxplot()+
  scale_x_discrete(labels = c("Better", "Worse", "Same", "Don't know"))+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    ylab("PC1")+
    xlab("Health Changes")

ggplot(subset(Metals_PCA, !is.na(Metals_PCA$Education)), aes(y=PC1, x=as.factor(Education)))+
    geom_boxplot()+
  scale_x_discrete(labels = c("None", "Elementary", "Middle", "High", "College", "Advanced", "Other"))+
    theme(legend.position = "none",
          panel.background = element_rect(fill="#ffffff"),
          axis.text.x=element_text(angle=60, hjust=1),
          axis.line = element_line(color="#000000"))+
    ylab("PC1")+
    xlab("Education")

```

