---
title: "Yuma Metals Empirical Distributions and Dimension Reduction - v3"
author: "Dean Billheimer"
date: "2021-08-28"
output:
html_document:
self_contained: yes
---

# Overview
Explore the distributional properties of (hair?) metals concentrations from
Yuma, AZ.  Also, use Rmd file to illustrate use of R and markdown for
computationally reproducible research.

### Changes since 2021-08-21
1. Values missing all metals are true missing -- insuffucent sample.
2. Values with missing Hg are also true missing (about 89 of these)
- fit PCA excluding missing Hg
- 2nd fit using imputed Hg from other metals data.
3. drop sample R1108 - missing Cu value - not identified in spectrum. This was included in a batch of
   samples that failed QA/QC check 

### Major changes since 2020-07-03 version
1. Updated data from Jonathan Credo - using 2021-08-23 version - sheet 3
2. Many of the NAs are true missing -
3. Jonathan re-coded BDL values to DL/sqrt(2)
4. Re-run the analysis.
5. Produce a new data set with robust PCA components for Jonathan's use.

### Major changes since 2020-06-23 version
1. Omit observations where most metals are unmeasured 
2. Omit Pb206 and 207
3. What to do about Hg NAs?

### Takeaway
Standard and robust PCA yield generally similar results.  There are enough
differences, however, that the robust method would be my first choice for
data summarization.

# R Stuff
## Set-up Environment and Read Data
```{r, setup}
library(readxl)
library(tidyverse)
library(ggplot2)

#setwd
setwd("~/Desktop/Research/Yuma/Yuma - R Project/")

## read data
##filename <- c("Yuma Metals Project - EMR 19Jun20.xlsx")
filename <- c("Yuma Metals Project_EMR_Editted 23Aug2021 mental health & ID removed.xlsx") ## new data set
tmp <- read_excel(filename, sheet=3)
dim(tmp) ## 330 222 ## same dimensions
 names(tmp)
 ##  [1] "ID"                        "Mn55"                     
 ##  [3] "Cu65"                      "Cd111"                    
 ##  [5] "Hg202"                     "Pb206"                    
 ##  [7] "Pb207"                     "PB208"                    
 ##  [9] "U238"                      "Site"                     
 ## [11] "Date"                      "Gender"                   

## head(tmp)

## keep only ID, metals and Site, cast to data frame
metals <- data.frame(tmp[,c(1:8)])
names(metals)  ## omitted Pb206 and Pb207
## metals check format
```

## Perform Initial Checks and Summary

The summary below shows there are 123 NAs for nearly all of the metals. Hg202 is
the exception with 212 NAs.

Data exploration shows that the NA pattern is consistent across the metals. If
one is missing, all are missing.

There are fewer measures at site 1.  More missing values at site 1 (51/78) and
at site 3 (48/126) compared with site 2 (24/126)

```{r, checks}
## Initial summary
summary(metals)

## Number of Observations per Site
table(metals$Site) ## 78 126 and 126.  Fewer at site 1

## Number of values > BDL by Site
table(metals$Site, !is.na(metals[,2])) ## Observation indicator by site
```



### Metals Concentrations
The figure below shows metals concentrations on log10 scale.  NA values are
omitted.

The shape of the Pb distributions are nearly identical (omitted 206 and 207). On the whole, these
distributions are not as bad as I feared.  All are stil right-skewed after
transformation, but they do not look pathologic.

```{r, intitialPlots, fig.cap="Metals concentrations - log10 scale"}
## change from wide to long format for plotting

metals %>% pivot_longer(-c(ID, Site), names_to="Metal", values_to="Concentration") %>%
    ggplot(aes(x=Concentration)) + geom_density() +
        scale_x_log10() + facet_wrap(~ Metal) +
            ggtitle("Distributions of Metals Concentrations (log10 transform)")
```

### Remove Missing Values

Values labeled as NA across (nearly all metals) are not measured.
For each observation with missing metals, omit the observation.  In addition
there are 89 observations with missing Hg values.  We may impute these missing values
based on site and other observed metals



*Note* - Other BDL values are replaced with 1/2 the limit of detection.

```{r, omitmissing}
names(metals)
metals.obs <- metals[!is.na(metals[,2]), ]
## for Hg, replace NA and zero with 0.029/sqrt(2)
##sort(metals.obs[,"Hg202"])
## Hg.subs <- 0.029 /sqrt(2)

metals.complete.cases <- metals.obs %>% filter(!is.na(Cu65)) %>% 
    filter(!is.na(Hg202)) ## ID R1108 also missing Hg, other metals observed

dim(metals.complete.cases) ## 111 8 wow - small data set
summary(metals.complete.cases)

metals.impute.Hg <- metals.obs %>% filter(!is.na(Cu65)) %>% 
    filter(is.na(Hg202))

metals.complete.log10 <- cbind(ID=metals.complete.cases[,"ID"],
                               log10(metals.complete.cases[,2:7]),
                               Site=metals.complete.cases[,"Site"])

summary(metals.complete.log10)
round(cor(metals.complete.log10[,2:7]), 2)

metals.impute.log10 <- cbind(ID=metals.impute.Hg[,"ID"],
                               log10(metals.impute.Hg[,2:7]),
                               Site=metals.impute.Hg[,"Site"])
summary(metals.impute.log10)
metals.impute.Hg

### fit regression line for Hg
Hg.lm <- lm(Hg202 ~ as.factor(Site) + Mn55 + Cu65 + Cd111 + Pb208 + U238,
            data=metals.complete.log10) 
qplot(metals.complete.log10[,"Hg202"], predict(Hg.lm))
qplot( predict(Hg.lm), resid(Hg.lm))

Hg.imputed <-  predict(Hg.lm, newdata=metals.impute.log10)

metals.fillin.log10 <- metals.impute.log10
metals.fillin.log10[,"Hg202"] <- Hg.imputed

summary(metals.fillin.log10)

## combine complete cases with Hg imputed cases
metals.allobs.log10 <- rbind(metals.complete.log10, metals.fillin.log10)
## dim(metals.allobs.log10) ## 206 8  ## OK obs with missing Cu65 is omitted.
## dim(metals.obs) ## 207 8

## check summary
summary(metals.allobs.log10)
```

### Adjusted (Hg Imputed) Concentrations
New figure with Hg values imputed using regression on other metals

```{r, checkPlots, fig.cap="Metals concentrations - log10 scale"}
## change from wide to long format for plotting

metals.allobs.log10 %>% pivot_longer(-c(ID, Site), names_to="Metal", values_to="Concentration") %>%
    ggplot(aes(x=Concentration)) + geom_density() +
         facet_wrap(~ Metal) +
            ggtitle("Distributions of Metals Concentrations (missin Hg values imputed, log10 transform)")
```

## Robust PCA

Now we move on to Robust PCA.  We use the method described by the following papers.

1. Hubert, M., Rousseeuw, P. J., and Vanden Branden, K. (2005), “ROBPCA: A New
   Approach to Robust Principal Component Analysis,” Technometrics, 47, 64–79. 

2. Engelen, S., Hubert, M. and Vanden Branden, K. (2005), “A Comparison of Three
   Procedures for Robust PCA in High Dimensions", Austrian Journal of
   Statistics, 34, 117–126. 

3. Hubert, M., Rousseeuw, P. J., and Verdonck, T. (2009), “Robust PCA for Skewed
   Data and Its Outlier Map," Computational Statistics & Data Analysis, 53,
   2264–2274. 

This is implemented in the R package $rospca$


The plot below shows results from the robust PCA.  The "Score distance"
describes the leverage that an indvidual point has in determining the fitted PCA
surface. The "Orthogonal distance" shows the distance to the fitted surface.

Thus, points 31, 48, 85, and 200 heaviy influence the slope of the PCA surface
(the loadings). Conversely, point 143 is a large distance from the mass of
points, but has little influence on the loadings.

The eigenvalues (bottom of the box) show that of the variation is captured
in PC1 (41%) and PC2 (20%).

```{r, robpca}
##install.packages("rospca", repos="http://cran.r-project.org")
library(rospca)

##dim(metals.adjusted)
##metals.log10 <- as.matrix(log10(metals.adjusted[, 2:7]))
colnames(metals.allobs.log10)

## round(cor(metals.log10), 2) ## check pairwise correlation
### == original full PCA

metals.pca <- robpca(metals.allobs.log10[,2:7], k=5, alpha=.7, ndir="all", skew=T)

## Eigenvalues show variation in different components
cat("Total Variance in PC1 and PC2\n")
round(metals.pca[[2]]/sum(metals.pca[[2]]), 3) ## variation in PC's is as expected across components

diagPlot(metals.pca)
```

The eigenvalues show that PC1 and PC2 exhibit the majority of the variance, PC
3-5 are smaller.  In general, the diagnostic plot suggests that robust PCA is warranted

### PCA Loadings
The loadings are shown below. 


```{r, loadings}
round(metals.pca$loadings, 3)
names(metals.allobs.log10[,2:7])
##plot(metals.pca$loadings, pch=rownames(metals.pca$loadings))
```
There are a few of points to notice about the loadings

1. All components of PC1 have the same sign. As before, this suggests a "common" effect to
   all measurements. For example, all metals tend to be "high" or "low"
   together. May be caused either by global high-low exposure of inviduals, or
   by mass spectrometry measurement issues.

2. PC1 has highest loading for Pb208, Mn55, and Cu65

2. For PC2, U238 is the biggest driver. 


### PCA Scores for Individuals

Finally, we look at PCA scores by site.  The second figure below shows clearly
that site 2 has slightly more variation than site 3. But not bad. It appears
that site 3 may be slightly greater on PC2 than is site 2.


```{r, pcaplot}
##names(metals.allobs.log10)
metals.pca.df <- data.frame(Site=factor(metals.allobs.log10$Site), metals.pca$scores)

qplot(x=PC1, y=PC2, data=metals.pca.df, col=Site,
      title="Robust PCA Scores by Site")

qplot(x=PC1, y=PC2, data=metals.pca.df) + ggtitle("Robust PCA Scores by Site") +
    facet_wrap(~ Site)


## dim(metals.pca$scores)
## dim(metals.allobs.log10)
## New df on 2021.09.06
metals.robpca <- cbind(metals.allobs.log10, metals.pca$scores)
## dim(metals.robpca)
### about here
qplot(x=PC1, y=Mn55, data=metals.robpca, col=as.factor(Site))
qplot(x=PC1, y=Pb208, data=metals.robpca, col=as.factor(Site))
qplot(x=PC2, y=U238, data=metals.robpca, col=as.factor(Site))


write.csv(metals.robpca, file="metals_Impute_log10_PCA_scores.csv", quote=F, row.names=F)

```

```{r, omitHg}
pca_No_Hg <- robpca(metals.allobs.log10[,c(2:4, 6, 7)], k=5, alpha=.7, ndir="all",
                    skew=T)
round(pca_No_Hg[[2]]/sum(pca_No_Hg[[2]]), 3) ## .48  .19 .11
round(pca_No_Hg$loadings, 2)
## dim(pca_No_Hg$scores) ## 206 5
metals.noHg.robpca <- cbind(metals.allobs.log10, pca_No_Hg$scores)

write.csv(metals.noHg.robpca, file="metals_NoHg_log10_PCA_scores.csv", quote=F, row.names=F)
```

```{r, render, eval=F}
##system("pwd")
rmarkdown::render("yuma-metals-20210821.Rmd", 'html_document')
```
